---
title: "DA220 Final Project"
author: "Ha Doan"
date: "3/24/2023"
output: html_document
---
## Purpose
- The purpose of this project to understand the relationship between genres and platforms to sales of video games around the world 

## Data Exploration

```{r}
library(tidyverse)
library(ggplot2)
library(tidymodels)
library(ggthemes)
library(tidytext)
library(wordcloud)
library(tm)
library(GGally)
library(ggfortify)
```

```{r}
myData <- read.csv("vgsales.csv")
myData <- subset(myData, select = -c(Rank), Year !="2020")  #drop the column Rank since it does not serve any purpose
# also eliminate ONE data point with the year 2020, as it might skew the trend later on
head(myData)
```




#### Add decade column for easier comparison/ trend
```{r}
myData <- myData %>%
  mutate(decade = as.numeric(Year)%/%10) 
```

```{r}
myData %>%
  select(decade, Publisher, Global_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(Global_Sales))
```

We have the overall sale statistic summary as following: 
```{r}
myData %>%
  summarise_if(is.numeric, funs(mean,sd,median,sum))
```


## Analysis and Model

### 1. Sales

#### 1.1 Sales over the decade:
```{r}
df1 <- myData %>%
  group_by(decade) %>%
  mutate(pctNa = mean(NA_Sales/Global_Sales * 100), 
         pctEu = mean(EU_Sales/Global_Sales * 100),
         pctJp = mean(JP_Sales/Global_Sales * 100),
         pctOther = mean(Other_Sales/Global_Sales * 100),
         avgGlobal = mean(Global_Sales))
df1
```




```{r}
ggplot(data=df1, aes(x=decade, y=pctNa, group=1)) +
  geom_line(color="maroon", size=1.25) +
  labs(title= "North American Sales Proportion over the Decades", x="Decade", y="Proportion of NA Sales (%)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), axis.text.x = element_text(angle=45), plot.caption = element_text(hjust=0.5, size=8))
```

```{r}
ggplot(data=df1, aes(x=decade, y=pctEu, group=1)) +
  geom_line(color="green3", size = 1.25) +
  labs(title= "European Sales Proportion over the Decades", x="Decade", y="Proportion of EU Sales (%)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), axis.text.x = element_text(angle=45), plot.caption = element_text(hjust=0.5, size=8))
```

```{r}
ggplot(data=df1, aes(x=decade, y=pctJp, group=1)) +
  geom_line(color="green3", size=1.25) +
  labs(title= "Japanese Sales Proportion over the Decades", x="Decade", y="Proportion of JP Sales (%)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), axis.text.x = element_text(angle=45), plot.caption = element_text(hjust=0.5, size=8))
```

```{r}
ggplot(data=df1, aes(x=decade, y=pctOther, group=1)) +
  geom_line(color="green3", size=1.25) +
  labs(title= "Other Region Sales Proportion over the Decades", x="Decade", y="Proportion of Other Sales (%)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), axis.text.x = element_text(angle=45), plot.caption = element_text(hjust=0.5, size=8))
```










######Inference for Linear Regression
Is the correlation between the two large enough

1) Hypotheses:

H0: rho = 0 (no lin rel) Ha: rho != 0 (yes, some relationship)

2) Test stats

sample size
```{r}
length(myData$JP_Sales)
```
df = 16597 - 2 = 16595
t value
```{r}
.61*sqrt(16597-2)/sqrt(1-(.61)^2)
```
3) p-value
```{r}
pt(99.17,16595)
```

##### Other Sales and EU Sales

```{r}
ggpairs(myData, columns=c("NA_Sales","EU_Sales","JP_Sales", "Other_Sales"))
```



```{r}
ggplot(myData, aes(EU_Sales, Other_Sales)) + 
  geom_point() + geom_smooth(method = "lm") + 
  labs(title= "Correlation Between European Sales vs Other Sales", x = "European Sales", y= "Other Sales") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), plot.caption = element_text(hjust=0.5, size=8))
```

###### Infernece for linear
Is the correlation between the two large enough

1) Hypotheses:

H0: rho = 0 (no lin rel) Ha: rho != 0 (yes, some relationship)

2) Test stats

df = 16597 - 2 = 16595

t-value
```{r}
.72*sqrt(16597-2)/sqrt(1-(.72)^2)
```
p-value
```{r}
1- pt(133.70,16595)
```
-> very small, approximately 0. Reject null hypothesis and there is some relationship between the 2

##### linear regression

Can we predict Other Sales based on EU?
```{r}
OtherModel <- lm(Other_Sales ~ EU_Sales, data = myData)
summary(OtherModel)
```


```{r}
autoplot(OtherModel)
```



### 2. Platform

To compare which platforms are driving Global Sales, we need calculate the its means by each platforms:
```{r}
myData %>%
  group_by(Platform) %>%
  summarise(total_charted = n()) %>%
  arrange(desc(total_charted))

```
```{r}
top5Platform <- myData %>%
          filter(Platform == "DS" | Platform == "PS2" | Platform == "PS3"|
                 Platform == "Wii"| Platform == "X360") 
```


```{r}
top5Platform %>%
  group_by(Platform) %>%
  summarise_if(is.numeric, funs(mean,sd)) %>%
  arrange(desc(Global_Sales_mean))  #arrange according to Global Sales, from top to bottom


```

We also have the following box plot to illustrate the distribution of the global sales across these 5 top platforms
```{r}
ggplot(top5Platform, aes(x = Platform, y = Global_Sales, fill = Platform)) +
  geom_boxplot() +
  labs(title= "Global Sales of Video Games in Top 5 Platforms", x="Platforms", y= "Global Sales") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), plot.caption = element_text(hjust=0.5, size=8))
```

Hard to see, a lot of outliers -> run a anova test

##### anova test for platforms

```{r}
anova_1 <- aov(top5Platform$Global_Sales ~ top5Platform$Platform)
summary(anova_1)
```
Strong evidence to suggest of a difference between mean Global Sales across these 5 platforms

##### multivariate linear regresison for platforms 

```{r}
globalPlatform <- lm(Global_Sales~Platform, data=top5Platform)
summary(globalPlatform)
```


```{r}
plot(globalPlatform$fitted.values, globalPlatform$residuals)
```


##### trend of platforms 
- Global
```{r}
trend1 <- top5Platform %>%
  select(decade, Platform, Global_Sales) %>%
  group_by(decade,Platform) %>%
  summarize(sale = mean(Global_Sales))

trend1 <- trend1 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

trend1
```


```{r}
ggplot(trend1, aes(x = decade, y = pct, group = Platform, color = Platform)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Global Sales of Top 5 Platforms by Decade",
       x = "Decades",
       y = "Global Sales (%)",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8)) 
```

- North America
```{r}
trend2 <- top5Platform %>%
  select(decade, Platform, NA_Sales) %>%
  group_by(decade,Platform) %>%
  summarize(sale = mean(NA_Sales))

trend2 <- trend2 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trend2, aes(x = decade, y = pct, group = Platform, color = Platform)) +
  geom_line() + geom_point(size=3) +
  labs(title = "North American Percentage Sales of Top 5 Platforms by Decade",
       x = "Decades",
       y = "Percentage Sales",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))

```

```{r}
trend3 <- top5Platform %>%
  select(decade, Platform, EU_Sales) %>%
  group_by(decade,Platform) %>%
  summarize(sale = mean(EU_Sales))

trend3 <- trend3 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trend3, aes(x = decade, y = pct, group = Platform, color = Platform)) +
  geom_line() + geom_point(size=3) +
  labs(title = "European Percentage Sales of Top Platforms by Decade",
       x = "Decades",
       y = "Percentage Sales",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

```{r}
trend4 <- top5Platform %>%
  select(decade, Platform, JP_Sales) %>%
  group_by(decade,Platform) %>%
  summarize(sale = mean(JP_Sales))

trend4 <- trend4 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trend4, aes(x = decade, y = pct, group = Platform, color = Platform)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Japanese Percentage Sales of Top Platforms by Decade",
       x = "Decades",
       y = "Percentage Sales",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```
```{r}
trend5 <- top5Platform %>%
  select(decade, Platform, Other_Sales) %>%
  group_by(decade,Platform) %>%
  summarize(sale = mean(Other_Sales))

trend5 <- trend5 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trend5, aes(x = decade, y = pct, group = Platform, color = Platform)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Other Countries Percentage Sales of Top Platforms by Decade",
       x = "Decades",
       y = "Percentage Sales",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```


### 3) Genres

```{r}
myData %>%
  group_by(Genre) %>%
  summarise(total_charted = n()) %>%
  arrange(desc(total_charted))
```
```{r}
top5Genre <- myData %>%
          filter(Genre == "Action" | Genre == "Sports" | Genre == "Misc" | 
                Genre == "Role-Playing" | Genre == "Shooter" ) 

head(top5Genre,10)
```


```{r}
ggplot(top5Genre, aes(x = Genre, y = Global_Sales, fill = Genre)) +
  geom_boxplot() +
  labs(title= "Global Sales of Video Games in Top 5 Genres", x="Genres", y= "Global Sales") +
  theme(plot.title = element_text(hjust = 0.5, size = 14), legend.title = element_text(size = 10), legend.text = element_text(size=8), axis.text = element_text(size=8), plot.caption = element_text(hjust=0.5, size=8))
```
Hard to see, a lot of outlier

#### ANOVA Test to see if there is difference in sales across genres


```{r}
anova_2 <- aov(top5Genre$Global_Sales ~ top5Genre$Genre)
summary(anova_2)
```
Yes there are. More specifically, look at the table below we could see that Shooter have the highest global sales
```{r}
top5Genre %>%
  group_by(Genre) %>%
  summarise_if(is.numeric, funs(mean,sd)) %>%
  arrange(desc(Global_Sales_mean))
```


#### multivariate linear regression 
Does that mean Genre could predict the sale of Global Sales?

```{r}
globalGenre <- lm(Global_Sales~Genre, data=top5Genre)
summary(globalGenre)
```
-> p-value > 0.05 in a lot of case, and adjusted R-squared is very small -> does not work. And so we could not predict Global Sales based on Genres (no predictive relationship between the two). Just because genre in top 5 does not automatically means higher global sales 

#### trend of genres across region
- Global:

```{r}
trendG1 <- top5Genre %>%
  select(decade, Genre, Global_Sales) %>%
  group_by(decade,Genre) %>%
  summarize(sale = mean(Global_Sales))

trendG1 <- trendG1 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendG1, aes(x = decade, y = pct, group = Genre, color = Genre)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Global Sales of Top 5 Genres across Decades",
       x = "Decades",
       y = "Global Sales (%)",
       color = "Genres") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

- North America:
```{r}
trendG2 <- top5Genre %>%
  select(decade, Genre, NA_Sales) %>%
  group_by(decade,Genre) %>%
  summarize(sale = mean(NA_Sales))

trendG2 <- trendG2 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendG2, aes(x = decade, y = pct, group = Genre, color = Genre)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of North American Sales of Top 5 Genres across Decades",
       x = "Decades",
       y = "NA Sales (%)",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

Have similar trend to global sales
- Europe:
```{r}
trendG3 <- top5Genre %>%
  select(decade, Genre, EU_Sales) %>%
  group_by(decade,Genre) %>%
  summarize(sale = mean(EU_Sales))

trendG3 <- trendG3 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendG1, aes(x = decade, y = pct, group = Genre, color = Genre)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of European Sales of Top 5 Genres across Decades",
       x = "Decades",
       y = "Percentage Sales",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

```{r}
trendG4 <- top5Genre %>%
  select(decade, Genre, JP_Sales) %>%
  group_by(decade,Genre) %>%
  summarize(sale = mean(JP_Sales))

trendG4 <- trendG4 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendG4, aes(x = decade, y = pct, group = Genre, color = Genre)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Japanese Sales of Top 5 Genres across Decades",
       x = "Decades",
       y = "JP Sales (%)",
       color = "Genres") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

- Very interesting, Role Playing seems to be most popular while shooter at the bottom -> contrast to trends so far

```{r}
trendG6 <- top5Genre %>%
  select(decade, Genre, Other_Sales) %>%
  group_by(decade,Genre) %>%
  summarize(sale = mean(Other_Sales))

trendG6 <- trendG6 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendG6, aes(x = decade, y = pct, group = Genre, color = Genre)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Other Sales of Top 5 Genres across Decades",
       x = "Decades",
       y = "Other Sales (%)",
       color = "Platform") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```
Role playing popular at first, but dropped significantly in 1990s and shooter game is on the rise. 




### 5) Publisher
#### chisq test between publisher and genres
```{r}
chisq.test(myData$Publisher, myData$Genre)
```
some relationship

#### top 5 publisher
```{r}
myData %>%
  group_by(Publisher) %>%
  summarise(total_charted = n()) %>%
  arrange(desc(total_charted))

```
```{r}
top5Publisher <- myData %>%
          filter(Publisher == "Electronic Arts" | Publisher == "Activision" | Publisher == "Namco Bandai Games" | Publisher== "Ubisoft" | Publisher == "Konami Digital Entertainment" ) 

head(top5Publisher,10)
```

```{r}
trendP1 <- top5Publisher %>%
  select(decade, Publisher, Global_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(Global_Sales))

trendP1 <- trendP1 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendP1, aes(x = decade, y = pct, group = Publisher, color = Publisher)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Global Sales of Top 5 Publishers across Decades",
       x = "Decades",
       y = "Global Sales (%)",
       color = "Publishers") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```
#### NA Publisher trend
```{r}
trendP2 <- top5Publisher %>%
  select(decade, Publisher, NA_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(NA_Sales))

trendP2 <- trendP2 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendP1, aes(x = decade, y = pct, group = Publisher, color = Publisher)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of North American Sales of Top 5 Publishers across Decades",
       x = "Decades",
       y = "NA Sales (%)",
       color = "Publishers") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```
#### Europe 
```{r}
trendP3 <- top5Publisher %>%
  select(decade, Publisher, EU_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(EU_Sales))

trendP3 <- trendP3 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendP3, aes(x = decade, y = pct, group = Publisher, color = Publisher)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of European Sales of Top 5 Publishers across Decades",
       x = "Decades",
       y = "European Sales (%)",
       color = "Publishers") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

#### JP sales
```{r}
trendP5 <- top5Publisher %>%
  select(decade, Publisher, JP_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(JP_Sales))

trendP5 <- trendP5 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendP5, aes(x = decade, y = pct, group = Publisher, color = Publisher)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Japanese Sales of Top 5 Publishers across Decades",
       x = "Decades",
       y = "Japanese Sales (%)",
       color = "Publishers") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

#### Other sales across publisher
```{r}
trendP6 <- top5Publisher %>%
  select(decade, Publisher, Other_Sales) %>%
  group_by(decade,Publisher) %>%
  summarize(sale = mean(Other_Sales))

trendP6 <- trendP6 %>%
  group_by(decade) %>%
  mutate(pct = sale/sum(sale) * 100)

ggplot(trendP6, aes(x = decade, y = pct, group = Publisher, color = Publisher)) +
  geom_line() + geom_point(size=3) +
  labs(title = "Proportion of Other Sales of Top 5 Publishers across Decades",
       x = "Decades",
       y = "Other Sales (%)",
       color = "Publishers") +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8))
```

#### percentage of genre per publisher
```{r}
ggplot(top5Publisher, aes(x = Publisher, fill = Genre)) +
  geom_bar(position = "fill") +
  labs(title = "Percentage of Genre per Publisher (in Top 5)",
       x = "Publisher",
       y = "Percentage",
       fill = "Publisher") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(plot.title = element_text(hjust = 0.5, size = 14),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 45),
        plot.caption = element_text(hjust = 0.5, size = 8)) +
  coord_flip()+
  scale_fill_manual(values=c("#FFA07A", "#F08080", "#FFB6C1", "#87CEEB", "#ADD8E6", "#66CDAA", "#3CB371", "#F0E68C", "#DAA520", "#BC8F8F", "#BDB76B", "#8FBC8F"))
```

